---
title: "EDS221 Day 3 Sessions"
format: html
editor: visual
---

## 

```{r}
temperature <- 97.2

if (temperature > 97) {
  print("Temperature warning")
} else 
    print("No warning")

```

Create an if else statement for blue whale length. Create a variable blue whale, which has a value of 40 meters. Create an if-else statement that returns "GIANT" if the length exceeds 35, "MEDIUM" if the length is between 20 and 35, or "SMALL" if the length is \<= 20

```{r}
blue_whale_length <- 40

if (blue_whale_length > 35) {
  print("GIANT")
} else if (20 <= blue_whale_length & blue_whale_length <= 35) {
  print("MEDIUM")
} else
  print("SMALL")
```

Let's say we are gonna be doing some experimental coding. I wanna try something out that isn't pushing the changes back to main without a review and making sure things are testable. So, I want to create... a branch!

We can do this through the git tab. It's those purple box thingees next to the main dropdown in the top right of RStudio.

In terminal: git checkout -B "name" --\> creates a new branch (-B) and switches over to it (checkout)

At this point, the branch is local. We need to make a remove version of the branch.

git push -u origin "name" --\> syncs the local branch with the remote branch

```{r}
housing_type <- "single"

switch(housing_type, 
       "single" = print("400"),
       "double" = print("750"),
       "whole" = print("1100"))
```
